---
title: "p8105_hw3_hc3654"
output: github_document
---

```{r}
library(p8105.datasets)
library(tidyverse)
data("instacart")
```


## Problem 1:

# Q1: How many aisles are there, and which aisles are the most items ordered from?

```{r}
p1q1 <- instacart |> 
  group_by(aisle) |> 
  summarise(Total_order = n()) |> 
  arrange(desc(Total_order))

p1q1
```
# Q2: Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
p1q2 <- 
p1q1 |> 
  filter(Total_order > 10000) |> 
  ggplot(aes(x = reorder(aisle,Total_order), y = Total_order,fill = Total_order))+
           geom_col()+
  coord_flip()+
  labs(
    x = "Aisle", y = "Total_order", title = "Figure shows the relates with aisle and their order among")
p1q2
  
  
```
# Q3: Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table

```{r}
p1q3 <- instacart |> 
  group_by(aisle,product_name) |> 
  summarise(product_sale = n()) |> 
  arrange(aisle, desc(product_sale)) |> 
  slice_head(n = 3)

p1q3
```

# Q4: Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
p1q4 <- instacart |> 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") |> 
  group_by(product_name, order_dow) |> 
  summarise(mean_hour = mean(order_hour_of_day)) |> 
  pivot_wider(
  names_from = order_dow,
  values_from = mean_hour
  ) 

           
knitr::kable(p1q4,
             col.names = c("Product","Sunday", "Monday", "Tuesday", "Wednesday", "Thursday","Friday", "Saturday"))

```

## Problem 2:

# 1.Import datasets

```{r}
ny_zillow_df <- 
  read_csv("zillow_data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv",na = c("NA", ".", " "))

zip_df <- 
  read_csv("zillow_data/Zip Codes.csv")

```

# Q1: There are 116 months between January 2015 and August 2024. How many ZIP codes are observed 116 times? How many are observed fewer than 10 times? Why are some ZIP codes are observed rarely and others observed in each month?

```{r}
p2 <- ny_zillow_df |> 
  janitor::clean_names() |> 
  pivot_longer(
    cols = x2015_01_31 : x2024_08_31,
    names_to = "date",
    values_to = "rental_fee"
  ) |> 
  mutate(date = sub("x", "",date))

p2q1 <-  p2 |> 
  group_by(region_name) |>
  summarise(n_months = sum(!is.na(rental_fee)), .groups = "drop")


obs116 <- p2q1 |> 
  filter(n_months == 116) |> 
  count()


obs10 <- p2q1 |> 
  filter(n_months < 10) |> 
  count()

p2q1 <- bind_cols(obs116,obs10)
knitr::kable(p2q1, col.names = c("Full months", "Less than 10"))
```

# Q2: Create a reader-friendly table showing the average rental price in each borough and year (not month). Comment on trends in this table.

```{r}
p2q2 <- p2 |> 
  separate(date, into = c("year", "month", "day"),sep = "_") |> 
  group_by(year, county_name) |> 
  summarise(avg_rental = mean(rental_fee, na.rm = TRUE)) 
p2q2
```

# Q3: Make a plot showing NYC Rental Prices within ZIP codes for all available years. Your plot should facilitate comparisons across boroughs. Comment on any significant elements of this plot.

```{r}

p2g1 <- 
p2 |> 
  group_by(region_name, date, county_name) |> 
  summarise(avg_price = mean(rental_fee, na.rm = TRUE), .groups = "drop") |> 
  mutate(date = ymd(date))|> 
  ggplot(aes(x = date, y = avg_price, group = region_name, color = avg_price)) +
  geom_line(alpha = 0.5)+
  scale_x_date(
  breaks = as.Date(c("2015-01-01", "2018-01-01", "2021-01-01", "2024-01-01")),
  labels = c("2015", "2019", "2021", "2024")
)+
  geom_smooth(aes(group = county_name), method = "loess", se = FALSE, size = 1.2, color = "orange", alpha = 0.6)+

  theme(
  axis.text.x = element_text(angle = 45, hjust = 1) 
)+
  facet_grid(~county_name)+
  labs( x= "Average Price", y = "Time", title = "Average Rental Price Over Time Separated by County" )

p2g1

```

# Q4: Compute the average rental price within each ZIP code over each month in 2023. Make a reader-friendly plot showing the distribution of ZIP-code-level rental prices across boroughs; put differently, your plot should facilitate the comparison of the distribution of average rental prices across boroughs. Comment on this plot

```{r}
p2g2 <- p2 |> 
  separate(date, into = c("year", "month", "day"),sep = "_") |> 
  filter(year == 2023) |> 
  group_by(region_name, month, county_name) |> 
  summarise(avg_price = mean(rental_fee, na.rm = TRUE)) |> 
  ggplot(aes(x = as.numeric(month), y = avg_price, group = region_name,color = avg_price))+
  geom_line()+
  geom_smooth(aes(group = county_name), method = "loess", se = FALSE, size = 1.2, color = "green", alpha = 0.6)+
  facet_grid(~ county_name)+
  scale_x_continuous(breaks = c(1,3,6,9,12))+
  labs(x = "Month", y = "Average Rental Fee", Title = "Relationship between Average Rental Fee and Month")

p2g2
```


# Q5: Combination
```{r}
library(patchwork)

combined <- p2g1 / p2g2 
if (!dir.exists("results")) dir.create("results")
ggsave("results/combined_plot.png", combined, width = 12, height = 10, dpi = 300)

```


## Problem 3:

```{r}

```








